Focal interaction for multi-parameter input variables

### 2023-06-14 paper/paper.tex
Leave 台大
Friday shutdown
Working on citations

\jd{I haven't looked at the Abstract for a long time; will get back to it. Moving it now to the main document file.}

@@

Plots of predicted values of an outcome against predictors (often called effect plots or prediction plots, generalized here as “outcome plots”) are often a useful way to summarize the results of a statistical model. These can be used to illustrate the shape of model response and associated uncertainties, and may be easier to interpret than tables of coefficient estimates \citep{brambor_understanding_2006, berry_improving_2012, leeper2017interpreting}.

@@

We endeavor here to make a conceptual distinction between two functions of outcome plots: prediction and effects. 
The primary difference lies in how we describe the uncertainties around the central estimate. 
If our goal is to \emph{predict} what we learn about the outcome variable by measuring the focal predictor, then we may want to capture a variety of sources of uncertainty, including that due to the intercept, focal and non-focal predictors, and random effects.
Conversely, if we wish to focus on the \emph{effect} of a focal predictor only, we might want to isolate uncertainty due to coefficients associated with that predictor.
If we follow this convention, we expect effects plots to generally have narrower confidence intervals (CIs) than prediction plots, since less uncertainty is taken into account.

@@

We discuss the idea of a model “center” and show how it simplifies thinking about outcome plot choices, particularly in the case of ordinary linear models. We also discuss biases that arise from non-linear link functions in generalized linear models, and how such biases can be addressed.

@@

\begin{figure}
\begin{center}
\includegraphics{eco.splash.Rout.pdf}
\end{center}
\caption{Outcome plots with effect- (dotted straight lines) and prediction-style (dashed curved lines) \CIs. The panel on the left anchors the effect-style intervals at the model center; the panel on the right has a specific anchor chosen for biological reasons. Anchor point does not affect prediction-style confidence intervals. Points show the fitted data; the large grey point shows the mean values of predictor and the response.}
\flab{splash}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics{eco.uniMulti.Rout.pdf}
%%
\caption{Outcome plots with effect- (dotted straight lines) and prediction-style (dashed curved lines). The panel on the left shows the results of a univariate model fit; the panel on the right shows a multivariate fit (and effect-style lines only, see text).}
\jd{Need to play with parameters or something; slopes should differ sharply.}
%%
\flab{uniMulti}
\end{center}
\end{figure}
%%

@@

\begin{figure}
\begin{center}
\includegraphics{eco.multiPK.defAnchor.centers.plot.Rout.pdf}
\end{center}
\caption{Outcome plots for a model with a non-focal interaction. Black lines show an effect plot (mean estimate and CIs) made using the model center (average of linear predictor variables) as a reference point. Blue lines show the same approach made by centering based on input variables; due to the interaction the central prediction line does not go through the center point.}
\flab{centers}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics{eco.quadComp.Rout.pdf}
\end{center}
\caption{Outcome plots for a model with a non-linear response to the focal variable. The left plot is anchored at the model center, which does not correspond to a possible value for the focal input variable (nitrogen). The right-hand plot is anchored at the mean value of the input variable (nitrogen). }
\flab{quadComp}
\end{figure}

>>>>

@@

Plots of predicted values of an outcome against predictors (often called effect plots or prediction plots, generalized here as “outcome plots”) are often a useful way to summarize the results of a statistical model. These can be used to illustrate the shape of model response and associated uncertainties, and may be easier to interpret than tables of coefficient estimates \citep{searle_population_1980, brambor_understanding_2006, schielzeth2010simple, berry_improving_2012, leeper2017interpreting}.

@@

We endeavor here to make a conceptual distinction between two functions of outcome plots: prediction and effects. 
The primary difference lies in how we describe the uncertainties around the central estimate. 
If our goal is to \emph{predict} what we learn about the outcome variable by measuring the focal predictor, then we may want to capture a variety of sources of uncertainty, including that due to the intercept, focal and non-focal predictors, and random effects. This is a common approach and is implemented as a default in many software packages for outcome plots \citep{fox2009effect, leeper2017package, lenth2018package,marginaleffects2023package}. 
Conversely, if we wish to focus on the \emph{effect} of a focal predictor only, we might want to isolate uncertainty due to coefficients associated with that predictor.
If we follow this convention, we expect effects plots to generally have narrower confidence intervals (CIs) than prediction plots, since less uncertainty is taken into account.

@@

We discuss the idea of a model “center” and show how it simplifies thinking about outcome plot choices, particularly in the case of ordinary linear models. We also discuss biases that arise from non-linear link functions in generalized linear models \citep{duursma_bias_2003,duursma2003bias,hanmer2013behind}, and how such biases can be addressed. The proposed approach and R software package will complement the existing ones, \citep{fox2009effect, leeper2017package, lenth2018package,marginaleffects2023package}, by providing a straightforward way to generate effects plots (in our sense), and correct for bias in generalized linear models.

@@

\begin{figure}
\begin{center}
\includegraphics{eco.jd.splash.Rout.pdf}
\end{center}
\caption{Outcome plots with effect- (dotted straight lines) and prediction-style (dashed curved lines) \CIs. The panel on the left anchors the effect-style intervals at the model center; the panel on the right has a specific anchor chosen for biological reasons. Anchor point does not affect prediction-style confidence intervals. Points show the fitted data; the large grey point shows the mean values of predictor and the response.}
\flab{splash}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics{eco.jd.uniMulti.Rout.pdf}
%%
\caption{Outcome plots with effect- (dotted straight lines) and prediction-style (dashed curved lines). The panel on the left shows the results of a univariate model fit; the panel on the right shows a multivariate fit (and effect-style lines only, see text).}
\jd{Need to play with parameters or something; slopes should differ sharply.}
%%
\flab{uniMulti}
\end{center}
\end{figure}
%%

@@

\begin{figure}
\begin{center}
\includegraphics{eco.jd.multiPK.defAnchor.centers.plot.Rout.pdf}
\end{center}
\caption{Outcome plots for a model with a non-focal interaction. Black lines show an effect plot (mean estimate and CIs) made using the model center (average of linear predictor variables) as a reference point. Blue lines show the same approach made by centering based on input variables; due to the interaction the central prediction line does not go through the center point.}
\flab{centers}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics{eco.jd.quadComp.Rout.pdf}
\end{center}
\caption{Outcome plots for a model with a non-linear response to the focal variable. The left plot is anchored at the model center, which does not correspond to a possible value for the focal input variable (nitrogen). The right-hand plot is anchored at the mean value of the input variable (nitrogen). }
\flab{quadComp}
\end{figure}
